---
title: "An Introduction to Data Analysis using Programming"
author: "Sumedh Girish"
output:
  pdf_document:
    toc: true
    toc_depth: 2
---

\pagebreak

# How to use this book

\pagebreak

\pagebreak

# Introduction to the R Programming Paradigm

- R, while being powerful lacks abstraction

- To avoid having to reinvent the wheel, we use packages in R,
  which is basically code written by other people to make our life simpler

- The most popular and widely used R package(s) come under the umbrella of the "tidyverse"

- It streamlines the process of dealing with data, standardizing
  a way to write R code for data analysis
  
```{r results="hide"}
library("tidyverse")
```

### Does working more hours make you smoke more?

  
## 1. Importing the data

- We use the readr library to read data in R

```{r}
# Reading from a csv file ../raw_data/ESS11.csv
filename <- "../raw_data/ESS11.csv"
datatibble <- read_csv(filename, show_col_types = FALSE)
print(datatibble)
```

- Notice how the information we want is just a subset of this data

## 2. Filtering the data we have

- We use the dplyr and tidyr library to do this

\pagebreak

```{r}
filtered_data <- datatibble |>
  select("cgtsmok", "wkhct", "wkhtot", "tporgwk") |>
  drop_na()
print(filtered_data)
```

- notice how data has a lot of invalid row values

```{r}
cleaned_data <- filtered_data |>
  filter(cgtsmok >= 1 & cgtsmok <= 6) |>
  filter(tporgwk >= 1 & tporgwk <= 6) |>
  filter(wkhct >= 0 & wkhct <= 168) |>
  filter(wkhtot >= 0 & wkhtot <= 168)
print(cleaned_data)
```

